---
title: "Data Sources and Cleaning"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 8,
  fig.asp = 0.6,
  out.width = "90%"
)

library(tidyverse)
library(knitr)
library(kableExtra)
```

# Data Sources

Our project integrates data from three primary sources to analyze the relationship between Alzheimer's disease prevalence, socioeconomic factors, and environmental quality among middle-aged and older adults.

## CDC Alzheimer’s Disease and Healthy Aging Data (2015-2022)
```{r cdc-table, echo=FALSE}
tibble(
  Attribute = c("Source", "Dataset", "Survey", "Years", "Geographic Coverage", "File Used", "URL"),
  Details = c(
    "Centers for Disease Control and Prevention (CDC)",
    "Alzheimer's Disease and Healthy Aging Data",
    "Behavioral Risk Factor Surveillance System (BRFSS)",
    "2015–2022",
    "59 U.S. states and territories",
    "2015-2022 Alzheimer Data.csv",
    "https://data.cdc.gov/Healthy-Aging/Alzheimer-s-Disease-and-Healthy-Aging-Data/hfr9-rurv/about_data")) |>
  kable() |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```











[U.S. Census American Community Survey (ACS)](https://data.census.gov)




The CDC Alzheimer's Disease and Healthy Aging Data provides state-level estimates of health indicators among middle-aged and older adults. 

    



Source: Centers for Disease Control and Prevention (CDC)

Description: We obtained data from the CDC’s Alzheimer’s Disease and Healthy Aging Data set. This dataset is derived from the Behavioral Risk Factor Surveillance System (BRFSS).

Scope: The raw dataset covers survey years 2015–2022 and includes data from 59 states and territories.

Focus: We specifically filtered for topics related to “Cognitive Decline” to capture self-reported cognitive problems among middle-aged and older adults.



Description: We obtained data from the CDC’s Alzheimer’s Disease and Healthy Aging Data set. This dataset is derived from the Behavioral Risk Factor Surveillance System (BRFSS).

Scope: The raw dataset covers survey years 2015–2022 and includes data from 59 states and territories.

Focus: We specifically filtered for topics related to “Cognitive Decline” to capture self-reported cognitive problems among middle-aged and older adults.















2. CDC Alzheimer's Disease and Healthy Aging Data Portal (2015-2022)
    - Variables of interest:
        - Prevalence rate
        - Cognitive decline for Alzheimer's-related outcomes
    - Usage in project: proxy measure for Alzheimer's prevalence 
    - URL: https://data.cdc.gov/Healthy-Aging/Alzheimer-s-Disease-and-Healthy-Aging-Data/hfr9-rurv/about_data
    
    
    
    
    
# Data Import and Cleaning