---
title: "P8105_final_Alzheimei"
date: "2025-11-27"
author: "bl3175, cw3747, lg3450, yw4662"
output: github_document
---

```{r}
library(tidyverse)
library(patchwork)
library(scales)
library(dplyr)


knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## Data Import

```{r}
raw_az_data <- read.csv("./2015-2022 Alzheimer Data.csv") |>
  janitor::clean_names()

raw_az_data <- raw_az_data[-nrow(raw_az_data), ]
```

```{r}
az_clean <- raw_az_data |>
  filter(
    !is.na(data_value),
    !is.na(year_start),
    !is.na(location_abbr)
  )
```

## Descriptive

```{r}

az_clean |>
  summarise(
    n_rows = n(),
    n_states = n_distinct(location_abbr),
    year_min = min(year_start),
    year_max = max(year_start),
    n_topics = n_distinct(topic),
    n_questions = n_distinct(question)
  )
```

We obtained data from the Centers for Disease Control and Prevention (CDC) Alzheimer’s Disease and Healthy Aging Data, which are derived from the Behavioral Risk Factor Surveillance System (BRFSS). After excluding records with missing key identifiers, our analytic dataset contained **32,616** observations from **59** states and territories, covering survey years **2015–2022**. The dataset includes **39** distinct topics and specific questions on health status and health-related behaviors among middle-aged and older adults.

For this study, we focused on BRFSS items whose topic was labeled **“cognitive decline”** or **“cognitive impairment”**. These items capture self-reported cognitive problems among middle-aged and older adults, and we treat them as Alzheimer’s-related cognitive indicators rather than clinically confirmed Alzheimer’s disease.

## Data considering Alzheimei

```{r}
library(stringr)

az_cog <- az_clean |>
  filter(
    str_detect(topic, regex("cognitive decline|cognitive impairment",
                            ignore_case = TRUE))
  )
```


### sex and age summary

```{r}
az_cog_summary_base <- az_cog |>
  mutate(
    age_group = case_when(
      stratification_category1 == "Age Group" ~ stratification1,
      TRUE ~ "Overall"
    ),
    sex_group = case_when(
      stratification_category2 == "Sex" ~ stratification2,
      TRUE ~ "Overall"
    )
  )
```

### state × age × sex table
```{r}
az_cog_state_age_sex <- az_cog_summary_base |>
  group_by(
    location_abbr, location_desc,
    age_group,
    sex_group
  ) |>
  summarise(
    n_obs = n(),
    mean_value = mean(data_value, na.rm = TRUE),
    .groups = "drop"
  ) |>
  arrange(location_abbr, age_group, sex_group)

az_cog_state_age_sex
```



