---
title: "P8105_final_Alzheimei"
date: "2025-11-27"
author: "bl3175, cw3747, lg3450, yw4662"
output: github_document
---

```{r}
library(tidyverse)
library(patchwork)
library(scales)
library(dplyr)


knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## Data Import

```{r}
raw_az_data <- read.csv("./2015-2022 Alzheimer Data.csv") |>
  janitor::clean_names()

raw_az_data <- raw_az_data[-nrow(raw_az_data), ]
```

```{r}
az_clean <- raw_az_data |>
  filter(
    !is.na(data_value),
    !is.na(year_start),
    !is.na(location_abbr)
  )
```

## Descriptive

```{r}

az_clean |>
  summarise(
    n_rows = n(),
    n_states = n_distinct(location_abbr),
    year_min = min(year_start),
    year_max = max(year_start),
    n_topics = n_distinct(topic),
    n_questions = n_distinct(question)
  )
```

We obtained data from the Centers for Disease Control and Prevention (CDC) Alzheimer’s Disease and Healthy Aging Data, which are derived from the Behavioral Risk Factor Surveillance System (BRFSS). After excluding records with missing key identifiers, our analytic dataset contained **32,616** observations from **59** states and territories, covering survey years **2015–2022**. The dataset includes **39** distinct topics and specific questions on health status and health-related behaviors among middle-aged and older adults.

For this study, we focused on BRFSS items whose topic was labeled **“cognitive decline”** or **“cognitive impairment”**. These items capture self-reported cognitive problems among middle-aged and older adults, and we treat them as Alzheimer’s-related cognitive indicators rather than clinically confirmed Alzheimer’s disease.

## Data considering Alzheimei

```{r}
library(stringr)

az_cog <- az_clean |>
  filter(
    str_detect(topic, regex("cognitive decline|cognitive impairment",
                            ignore_case = TRUE))
  )
```


### sex and age summary

```{r}
az_cog_summary_base <- az_cog |>
  mutate(
    age_group = case_when(
      stratification_category1 == "Age Group" ~ stratification1,
      TRUE ~ "Overall"
    ),
    sex_group = case_when(
      stratification_category2 == "Sex" ~ stratification2,
      TRUE ~ "Overall"
    )
  )
```

### state × age × sex table
```{r}
az_cog_state_age_sex <- az_cog_summary_base |>
  group_by(
    location_abbr, location_desc,
    age_group,
    sex_group
  ) |>
  summarise(
    n_obs = n(),
    mean_value = mean(data_value, na.rm = TRUE),
    .groups = "drop"
  ) |>
  arrange(desc(mean_value))

az_cog_state_age_sex
```

### top 10 states (2015-2016)

```{r}
az_cog_65_overall <- az_cog_summary_base |>
  filter(
    age_group == "65 years or older",
    sex_group == "Overall"
  )

nrow(az_cog_65_overall)
sort(unique(az_cog_65_overall$year_start))
```

The CDC Healthy Aging dataset covers survey years 2015–2022. For topics coded as Cognitive Decline or Cognitive Impairment, data were only available in 2015–2016; therefore, our Alzheimer’s-related analyses are restricted to these two years.

```{r}
library(ggplot2)

az_cog_65_overall <- az_cog_summary_base |>
  filter(
    age_group == "65 years or older",
    sex_group == "Overall"
  )

years_use <- c(2015, 2016)

az_cog_65_overall_15_16 <- az_cog_65_overall |>
  filter(year_start %in% years_use)

```


```{r}
top10_state_15_16 <- az_cog_65_overall_15_16 |>
  group_by(location_abbr, location_desc) |>
  summarise(
    n_obs = n(),
    mean_15_16 = mean(data_value, na.rm = TRUE),
    .groups = "drop"
  ) |>
  arrange(desc(mean_15_16)) |>
  slice_head(n = 10)

top10_state_15_16
```

Using data from 2015–2016, we ranked states and regions by the mean prevalence of Alzheimer’s-related cognitive indicators among adults aged 65 years and older (all sexes combined). The highest burden was observed in Puerto Rico (mean prevalence 47.9%), which was notably higher than any U.S. state or region. The Northeast region ranked second (33.6%), followed by Oklahoma (31.5%), Mississippi (31.4%), and the South region as a whole (31.0%). The remaining states in the top ten—South Dakota, Massachusetts, Alabama, Alaska, and South Carolina—all had mean prevalences around 30%. Overall, these results suggest substantial geographic variation in self-reported cognitive problems among older adults, with several Southern states and Puerto Rico showing particularly high levels.

The Top 10 list for 2015–2016 is different from the earlier state–age–sex table because they summarize different things. The state–age–sex table ranks specific subgroups (for example, women aged 50–64 in Puerto Rico), so it shows where the highest subgroup prevalences occur. In contrast, the 2015–2016 Top 10 is based only on adults aged 65+ with sexes combined, giving one value per state. This ranking reflects overall burden in older adults, not the maximum value in any age–sex subgroup, and provides useful context for later stratified analyses or adjustment for potential confounders such as age, sex, or region.

## Distribution Difference

### prevelance ~ year

```{r}
az_nat_65_overall <- az_cog_65_overall |>
  group_by(year = year_start) |>
  summarise(
    mean_prev = mean(data_value, na.rm = TRUE),
    sd_prev   = sd(data_value, na.rm = TRUE),
    .groups = "drop"
  )

az_nat_65_overall |>
  mutate(year = factor(year)) |>
  ggplot(aes(x = year, y = mean_prev)) +
  geom_col() +
  labs(
    x = "Year",
    y = "Mean prevalence (%)",
    title = "Alzheimer’s-related cognitive indicators among adults 65+, 2015–2016"
  )
```

At the national level, the mean prevalence of Alzheimer’s-related cognitive indicators among adults aged 65+ was very similar in 2015 (26.0%) and 2016 (25.8%). The small difference of 0.2 percentage points is well within the between-state variation and does not suggest any clear change over this two-year period.

### prevelance ~ age

```{r}
az_age_desc <- az_cog_summary_base |>
  group_by(age_group) |>
  summarise(
    n_obs      = n(),
    mean_value = mean(data_value, na.rm = TRUE),
    .groups = "drop"
  )

ggplot(az_age_desc,
       aes(x = age_group, y = mean_value)) +
  geom_col() +
  labs(
    x = "Age group",
    y = "Mean prevalence (%)",
    title = "Alzheimer’s-related cognitive indicators by age group"
  )
```

Across all cognitive items, the mean prevalence was about 31.8% among adults aged 50–64 years and 25.8% among those aged 65 years or older.

### prevelance ~ sex

```{r}
az_sex_desc <- az_cog_summary_base |>
  group_by(sex_group) |>
  summarise(
    n_obs      = n(),
    mean_value = mean(data_value, na.rm = TRUE),
    .groups = "drop"
  )

ggplot(az_sex_desc,
       aes(x = sex_group, y = mean_value)) +
  geom_col() +
  labs(
    x = "Sex",
    y = "Mean prevalence (%)",
    title = "Alzheimer’s-related cognitive indicators by sex"
  )
```

Across all states and cognitive items in 2015–2016, women had a higher mean prevalence of Alzheimer’s-related cognitive indicators than men (30.9% vs 26.8%), suggesting a consistently greater reported cognitive burden among female respondents.



