---
title: "P8105_final_Alzheimei"
date: "2025-11-27"
author: "bl3175, cw3747, lg3450, yw4662"
output: github_document
---

```{r}
library(tidyverse)
library(patchwork)
library(scales)
library(dplyr)


knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## Data Import

```{r}
raw_az_data <- read.csv("data/2015-2022 Alzheimer Data.csv") |>
  janitor::clean_names()

raw_az_data <- raw_az_data[-nrow(raw_az_data), ]
```

```{r}
az_clean <- raw_az_data |>
  filter(
    !is.na(data_value),
    !is.na(year_start),
    !is.na(location_abbr)
  )
```

## Descriptive

```{r}

az_clean |>
  summarise(
    n_rows = n(),
    n_states = n_distinct(location_abbr),
    year_min = min(year_start),
    year_max = max(year_start),
    n_topics = n_distinct(topic),
    n_questions = n_distinct(question)
  )
```

We obtained data from the Centers for Disease Control and Prevention (CDC) Alzheimer’s Disease and Healthy Aging Data, which are derived from the Behavioral Risk Factor Surveillance System (BRFSS). After excluding records with missing key identifiers, our analytic dataset contained **32,616** observations from **59** states and territories, covering survey years **2015–2022**. The dataset includes **39** distinct topics and specific questions on health status and health-related behaviors among middle-aged and older adults.

For this study, we focused on BRFSS items whose topic was labeled **“cognitive decline”** or **“cognitive impairment”**. These items capture self-reported cognitive problems among middle-aged and older adults, and we treat them as Alzheimer’s-related cognitive indicators rather than clinically confirmed Alzheimer’s disease.

## Data considering Alzheimei

```{r}
library(stringr)

az_cog <- az_clean |>
  filter(
    str_detect(topic, regex("cognitive decline|cognitive impairment",
                            ignore_case = TRUE))
  )
```


### sex and age summary

```{r}
az_cog_summary_base <- az_cog |>
  mutate(
    age_group = case_when(
      stratification_category1 == "Age Group" ~ stratification1,
      TRUE ~ "Overall"
    ),
    sex_group = case_when(
      stratification_category2 == "Sex" ~ stratification2,
      TRUE ~ "Overall"
    )
  )
```

### state × age × sex table
```{r}
az_cog_state_age_sex <- az_cog_summary_base |>
  group_by(
    location_abbr, location_desc,
    age_group,
    sex_group
  ) |>
  summarise(
    n_obs = n(),
    mean_value = mean(data_value, na.rm = TRUE),
    .groups = "drop"
  ) |>
  arrange(desc(mean_value))

az_cog_state_age_sex
```

### top 10 states (2015-2016)

```{r}
az_cog_65_overall <- az_cog_summary_base |>
  filter(
    age_group == "65 years or older",
    sex_group == "Overall"
  )

nrow(az_cog_65_overall)
sort(unique(az_cog_65_overall$year_start))
```

The CDC Healthy Aging dataset covers survey years 2015–2022. For topics coded as Cognitive Decline or Cognitive Impairment, data were only available in 2015–2016; therefore, our Alzheimer’s-related analyses are restricted to these two years.

```{r}
library(ggplot2)

az_cog_65_overall <- az_cog_summary_base |>
  filter(
    age_group == "65 years or older",
    sex_group == "Overall"
  )

years_use <- c(2015, 2016)

az_cog_65_overall_15_16 <- az_cog_65_overall |>
  filter(year_start %in% years_use)

```


```{r}
top10_state_15_16 <- az_cog_65_overall_15_16 |>
  group_by(location_abbr, location_desc) |>
  summarise(
    n_obs = n(),
    mean_15_16 = mean(data_value, na.rm = TRUE),
    .groups = "drop"
  ) |>
  arrange(desc(mean_15_16)) |>
  slice_head(n = 10)

top10_state_15_16
```

Using data from 2015–2016, we ranked states and regions by the mean prevalence of Alzheimer’s-related cognitive indicators among adults aged 65 years and older (all sexes combined). The highest burden was observed in Puerto Rico (mean prevalence 47.9%), which was notably higher than any U.S. state or region. The Northeast region ranked second (33.6%), followed by Oklahoma (31.5%), Mississippi (31.4%), and the South region as a whole (31.0%). The remaining states in the top ten—South Dakota, Massachusetts, Alabama, Alaska, and South Carolina—all had mean prevalences around 30%. Overall, these results suggest substantial geographic variation in self-reported cognitive problems among older adults, with several Southern states and Puerto Rico showing particularly high levels.

The Top 10 list for 2015–2016 is different from the earlier state–age–sex table because they summarize different things. The state–age–sex table ranks specific subgroups (for example, women aged 50–64 in Puerto Rico), so it shows where the highest subgroup prevalences occur. In contrast, the 2015–2016 Top 10 is based only on adults aged 65+ with sexes combined, giving one value per state. This ranking reflects overall burden in older adults, not the maximum value in any age–sex subgroup, and provides useful context for later stratified analyses or adjustment for potential confounders such as age, sex, or region.

## Distribution Difference

### prevelance ~ year

```{r}
az_nat_65_overall <- az_cog_65_overall |>
  group_by(year = year_start) |>
  summarise(
    mean_prev = mean(data_value, na.rm = TRUE),
    sd_prev   = sd(data_value, na.rm = TRUE),
    .groups = "drop"
  )

az_nat_65_overall |>
  mutate(year = factor(year)) |>
  ggplot(aes(x = year, y = mean_prev)) +
  geom_col() +
  labs(
    x = "Year",
    y = "Mean prevalence (%)",
    title = "Alzheimer’s-related cognitive indicators among adults 65+, 2015–2016"
  )
```

At the national level, the mean prevalence of Alzheimer’s-related cognitive indicators among adults aged 65+ was very similar in 2015 (26.0%) and 2016 (25.8%). The small difference of 0.2 percentage points is well within the between-state variation and does not suggest any clear change over this two-year period.

### prevelance ~ age

```{r}
az_age_desc <- az_cog_summary_base |>
  group_by(age_group) |>
  summarise(
    n_obs      = n(),
    mean_value = mean(data_value, na.rm = TRUE),
    .groups = "drop"
  )

ggplot(az_age_desc,
       aes(x = age_group, y = mean_value)) +
  geom_col() +
  labs(
    x = "Age group",
    y = "Mean prevalence (%)",
    title = "Alzheimer’s-related cognitive indicators by age group"
  )
```

Across all cognitive items, the mean prevalence was about 31.8% among adults aged 50–64 years and 25.8% among those aged 65 years or older.

### prevelance ~ sex

```{r}
az_sex_desc <- az_cog_summary_base |>
  group_by(sex_group) |>
  summarise(
    n_obs      = n(),
    mean_value = mean(data_value, na.rm = TRUE),
    .groups = "drop"
  )

ggplot(az_sex_desc,
       aes(x = sex_group, y = mean_value)) +
  geom_col() +
  labs(
    x = "Sex",
    y = "Mean prevalence (%)",
    title = "Alzheimer’s-related cognitive indicators by sex"
  )
```

Across all states and cognitive items in 2015–2016, women had a higher mean prevalence of Alzheimer’s-related cognitive indicators than men (30.9% vs 26.8%), suggesting a consistently greater reported cognitive burden among female respondents.


## Income Effects

### data import

```{r}
inc15 <- read_csv("data/median income-2015.csv", skip = 1) |>
  janitor::clean_names() |>
  transmute(
    year = 2015,
    state_name = geographic_area_name,
    median_income = as.numeric(median_income_dollars_estimate_households)
  )


inc16 <- read_csv("data/median income-2016.csv", skip = 1) |>
  janitor::clean_names() |>
  transmute(
    year = 2016,
    state_name = geographic_area_name,
    median_income = as.numeric(median_income_dollars_estimate_households)
  )

income_state <- bind_rows(inc15, inc16)
```

### income by state

```{r}
income_state_with_state <- income_state |>
  mutate(
    state = str_squish(str_extract(state_name, "[^,]+$")))

income_state_level <- income_state_with_state |>
  group_by(year, state) |>
  summarise(
    n_counties = n(),
    median_income_state = median(median_income, na.rm = TRUE),
    .groups = "drop"
  )

income_state_level <- income_state_with_state |>
  group_by(year, state) |>
  summarise(
    n_counties = n(),
    median_state = median(median_income, na.rm = TRUE),
    mean_state   = mean(median_income, na.rm = TRUE),
    .groups = "drop"
  )

income_state_level |> head(10)
```


As the Healthy Aging dataset is designed to monitor cognitive health in older adults and Alzheimer’s disease primarily affects seniors, we restricted our analytic sample to respondents aged 65+ years. For the main state-level analysis, we used sex-grouped estimates labeled “Overall” rather than sex-specific values so that each state–year contributed a single prevalence estimate, simplifying the comparison with state-level income and avoiding multiple correlated observations per state.

```{r}
state_vec <- unique(income_state_level$state)

az_cog_65_overall_state <- az_cog_65_overall |>
  filter(
    year_start %in% c(2015, 2016),
    sex_group == "Overall",
    location_desc %in% state_vec
  )

az_income <- az_cog_65_overall_state |>
  left_join(
    income_state_level,
    by = c("year_start" = "year",
           "location_desc" = "state")
  )

```

### relationship

```{r}
ggplot(az_income,
       aes(x = median_state, y = data_value,
           color = factor(year_start))) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    x = "State-level median household income (USD)",
    y = "Alzheimer’s-related cognitive prevalence (%)",
    color = "Year",
    title = "State-level cognitive burden vs income among adults 65+ (2015–2016)"
  )
```

```{r}
model_inc <- lm(
  data_value ~ median_state + factor(year_start),
  data = az_income
)

summary(model_inc)
```

In a linear regression of state-level Alzheimer’s-related cognitive prevalence on median household income and survey year, higher income was associated with lower cognitive burden. The estimated coefficient for income was
$\hat\beta_{\text{income}} = -2.35 \times 10^{-4}$, meaning that a \$1000 increase in state median income was associated with about a \textbf{0.24–percentage-point lower} prevalence, and a \$10000 increase was associated with about a \textbf{2.35–percentage-point lower} prevalence among adults aged 65+. The coefficient for survey year (2016 vs 2015) was small ($\hat\beta = -0.29$), suggesting no clear overall difference between the two years after accounting for income. However, the model $R^2$ was only \textbf{0.04}, indicating that income explains only a small fraction of the large between-state variation in cognitive outcomes.

Although higher state-level income was statistically associated with lower Alzheimer’s-related cognitive prevalence, income explained only a small share of the between-state variability. Several factors may contribute to this weak relationship. First, the analysis is ecological: each observation is an aggregated state–year estimate, so large within-state differences in income and cognitive health are not captured, and any individual-level association is diluted. Second, both cognitive burden and income are measured with error (self-reported survey indicators and ACS estimates), and we further aggregated county-level income up to the state level, which may introduce additional noise. Third, we only examined two years (2015–2016), over which variation in both income and prevalence is fairly modest. Finally, many other state-level determinants—such as education, health care access, demographic composition, and prevalence of vascular risk factors—were not adjusted for, so income alone is unlikely to account for the wide variation in cognitive outcomes across states.


## Education Effects

### data import

```{r}
edu15_raw <- read_csv("data/education2015.csv", skip = 1) |>
  janitor::clean_names()

edu15_state <- edu15_raw |>
  filter(!str_detect(geographic_area_name, ","))

edu16_raw <- read_csv("data/education2016.csv", skip = 1) |>
  janitor::clean_names()

edu16_state <- edu16_raw |>
  filter(!str_detect(geographic_area_name, ","))
```

We used the proportion of adults aged 25 years and older with a bachelor’s degree or higher as our main education indicator. This measure is commonly reported in the ACS and captures individuals who have completed at least a four-year college education, which is often used as a marker of higher socioeconomic status and cognitive reserve in the literature. Compared with more detailed education categories, using “bachelor’s degree or higher” gives a simple, comparable summary of educational attainment across states and years.

```{r}
edu15_state_slim <- edu15_state |>
  transmute(
    year = 2015,
    state = geographic_area_name,
    educ_bach = as.numeric(percent_estimate_percent_bachelors_degree_or_higher)
  )

edu16_state_slim <- edu16_state |>
  transmute(
    year = 2016,
    state = geographic_area_name,
    educ_bach = as.numeric(percent_estimate_percent_bachelors_degree_or_higher)
  )

edu_state <- bind_rows(edu15_state_slim, edu16_state_slim)
```

```{r}
az_edu <- az_cog_65_overall_state |>
  left_join(
    edu_state,
    by = c(
      "year_start" = "year",
      "location_desc" = "state"
    ))
```

### relationship analysis

```{r}
ggplot(az_edu,
       aes(x = educ_bach,
           y = data_value,
           color = factor(year_start))) +
  geom_point(alpha = 0.8) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    x = "State-level % with bachelor’s degree or higher (25+)",
    y = "Alzheimer’s-related cognitive prevalence (%)",
    color = "Year",
    title = "State-level cognitive burden vs education among adults 65+ (2015–2016)"
  ) +
  theme_minimal()
```

```{r}
model_edu <- 
  lm(
  data_value ~ educ_bach + factor(year_start),
  data = az_edu)

summary(model_edu)
```


In a linear regression of state-level Alzheimer’s-related cognitive prevalence on the percentage of adults with a bachelor’s degree or higher and survey year, higher education was associated with lower cognitive burden. The estimated coefficient for education was $\hat\beta_{\text{educ}} = -0.16$, meaning that a 1–percentage-point increase in the share of adults 25 years and older with at least a bachelor’s degree was associated with about a \textbf{0.16–percentage-point lower} prevalence of Alzheimer’s-related cognitive problems among adults aged 65+. A 10–percentage-point increase in educational attainment would correspond to roughly a \textbf{1.6–percentage-point lower} prevalence. However, this association was relatively weak and not conventionally statistically significant (the approximate 95\% confidence interval for $\hat\beta_{\text{educ}}$ ranges from about $-0.35$ to $0.03$). The coefficient for survey year (2016 vs 2015) was small ($\hat\beta = -1.14$), suggesting no clear overall difference between the two years after accounting for education. The model $R^2$ was only about \textbf{0.01}, indicating that education and year together explain only a small fraction of the large between-state variation in cognitive outcomes.

Although states with higher levels of college education tended to have slightly lower Alzheimer’s-related cognitive prevalence, the relationship was weak and most of the between-state variability remained unexplained. Several factors may help explain this. First, the analysis is ecological: each data point is a state–year average, so big within-state differences in education and cognitive health are smoothed out, and any individual-level association is likely diluted. Second, both education and cognitive burden come from survey-based estimates and can be measured with error; on top of that, we aggregated county-level education indicators to the state level, which probably adds extra noise. Third, we only looked at two survey years (2015–2016), so there is not much variation over time in either education or cognitive outcomes. Finally, many other state-level factors—such as income, access to health care, population structure, and the prevalence of vascular and lifestyle risk factors—were not included in the model, so education alone is unlikely to explain the wide differences in cognitive outcomes across states.

## Alzheimer’s outcomes at state-year level

```{r}
az_state_year <- az_cog_65_overall |>
  filter(
    year_start %in% c(2015, 2016),
    sex_group == "Overall"
  ) |>
  group_by(
    year       = year_start,
    state_abbr = location_abbr,
    state_name = location_desc
  ) |>
  summarise(
    mean_prev = mean(data_value, na.rm = TRUE),
    .groups = "drop"
  )

head(az_state_year)
```

## Air Pollution Effects

### Data import

```{r}
air15 <- read_csv("data/clean_air_state_2015.csv")
air16 <- read_csv("data/clean_air_state_2016.csv")

air_state <- bind_rows(air15, air16)

air_state |> head()
```
We used state-level fine particulate matter (PM2.5) concentrations as our primary indicator of air pollution. Specifically, we relied on the 24-hour weighted mean PM2.5 values derived from EPA AirData annual summary reports. These data were originally reported at the county level and then aggregated to obtain state–year averages for 2015 and 2016, matching the years available in the CDC Healthy Aging cognitive dataset.

```{r}
air_state_level <- air_state |>
  group_by(year, state_abbr) |>
  summarise(
    pm25_state = mean(pm25_wtd_mean, na.rm = TRUE),
    .groups = "drop"
  )

air_state_level |> head(10)
```


### merge with Alzheimer’s outcomes

```{r}
az_cog_65_overall_state <- az_cog_65_overall |>
  filter(
    year_start %in% c(2015, 2016),
    sex_group == "Overall"
  )

az_air <- az_cog_65_overall_state |>
  left_join(
    air_state_level,
    by = c(
      "year_start"    = "year",
      "location_abbr" = "state_abbr"
    )
  )

az_air |>
  select(year_start, location_desc, pm25_state, data_value) |>
  head(10)
```

For consistency with the income and education analyses, we restricted the sample to adults aged 65+ years and used the sex-grouped estimates labeled “Overall”. Each state–year therefore contributes a single Alzheimer’s-related prevalence value, which we then merged with the corresponding state–year PM2.5 concentration.

### relationship analysis
```{r}
ggplot(az_air,
       aes(x = pm25_state,
           y = data_value, 
           color = factor(year_start))) +
  geom_point(alpha = 0.8) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    x = "State-level PM2.5 weighted mean (µg/m³)",
    y = "Alzheimer’s-related cognitive prevalence (%)",
    color = "Year",
    title = "State-level cognitive burden vs PM2.5 among adults 65+ (2015–2016)"
  ) +
  theme_minimal()
```

```{r}
model_air <- lm(
  data_value ~ pm25_state + factor(year_start),
  data = az_air
)

summary(model_air)
```

In a linear regression of state-level Alzheimer’s-related cognitive prevalence on PM2.5 and survey year, we did not observe a clear association between air pollution and cognitive burden among adults aged 65+. The estimated coefficient for PM2.5 was $\hat\beta_{\text{PM2.5}} =$ `r round(coef(model_air)[2], 3)`, meaning that a 1–µg/m³ increase in the state-level PM2.5 weighted mean was associated with an estimated `r round(coef(model_air)[2], 3)`–percentage-point change in Alzheimer’s-related cognitive prevalence. A 10–µg/m³ difference in PM2.5 would therefore correspond to roughly `r round(coef(model_air)[2] * 10, 2)` percentage points. However, this association was small and not statistically significant (p = `r round(summary(model_air)$coefficients[2, 4], 3)`), indicating that PM2.5 is not a strong predictor of between-state differences in cognitive burden in this model.

The coefficient for survey year (2016 vs 2015) was $\hat\beta_{\text{year}} =$ `r round(coef(model_air)[3], 3)` with p = `r round(summary(model_air)$coefficients[3, 4], 3)`, suggesting no meaningful overall change in prevalence between the two years after accounting for PM2.5. The model $R^2$ was only `r round(summary(model_air)$r.squared, 3)`, implying that PM2.5 and year together explain only a negligible fraction of the substantial between-state variation in Alzheimer’s-related cognitive outcomes. This is not surprising given the ecological nature of the analysis (state-level averages), potential measurement error in both pollution and cognitive indicators, the short two-year time window, and the absence of other important determinants such as income, education, demographics, and health care access.
	
## Mapping Alzheimer’s-related Cognitive Burden

### State-level choropleth map for 2015

```{r map_2015, message = FALSE, warning = FALSE}
library(maps)
library(stringr)

az_2015 = az_state_year |> 
  filter(year == 2015)

us_map = map_data("state") |> as_tibble()

az_map_2015 = az_2015 |> 
  mutate(region = str_to_lower(state_name)) |> 
  inner_join(us_map, by = "region")

ggplot(az_map_2015,
       aes(long, lat, group = group, fill = mean_prev)) +
  geom_polygon(color = "white", linewidth = 0.2) +
  coord_fixed(1.3) +
  scale_fill_viridis_c(option = "C") +
  labs(
    title = "Alzheimer’s-related cognitive prevalence (Age 65+, 2015)",
    fill  = "Prevalence (%)"
  ) +
  theme_void() +
  theme(legend.position = "right")
```

### State-level choropleth map for 2016

```{r map_2016, message = FALSE, warning = FALSE}
library(maps)
library(stringr)

az_2016 = az_state_year |> 
  filter(year == 2016)

us_map = map_data("state") |> as_tibble()

az_map_2016 = az_2016 |> 
  mutate(region = str_to_lower(state_name)) |> 
  inner_join(us_map, by = "region")

ggplot(az_map_2016,
       aes(long, lat, group = group, fill = mean_prev)) +
  geom_polygon(color = "white", linewidth = 0.2) +
  coord_fixed(1.3) +
  scale_fill_viridis_c(option = "C") +
  labs(
    title = "Alzheimer’s-related cognitive prevalence (Age 65+, 2016)",
    fill  = "Prevalence (%)"
  ) +
  theme_void() +
  theme(legend.position = "right")
```

### Side-by-side Comparison (2015 vs 2016)

```{r map_compare_2015_2016, message = FALSE, warning = FALSE}
az_15_16 = az_state_year |> 
  filter(year %in% c(2015, 2016))

us_map = map_data("state") |> as_tibble()

az_map_15_16 = az_15_16 |> 
  mutate(region = str_to_lower(state_name)) |> 
  inner_join(us_map, by = "region")

ggplot(az_map_15_16,
       aes(long, lat, group = group, fill = mean_prev)) +
  geom_polygon(color = "white", linewidth = 0.2) +
  coord_fixed(1.3) +
  scale_fill_viridis_c(option = "C") +
  labs(
    title = "Alzheimer’s-related cognitive prevalence (Age 65+, 2015–2016)",
    fill  = "Prevalence (%)"
  ) +
  theme_void() +
  theme(legend.position = "right") +
  facet_wrap(~ year)
```

Geographic patterns of Alzheimer’s-related cognitive prevalence were visualized using state-level choropleth maps for 2015 and 2016. In both years, several Southern and Southeastern states showed higher prevalence levels, while many Western and Northeastern states displayed noticeably lower burdens. The spatial distribution remained largely consistent between 2015 and 2016, suggesting limited short-term change in geographic patterns of cognitive burden among adults aged 65+.
