---
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    theme: flatly
---

<style>
/* Global Styles */
html, body {
  font-family: Georgia, 'Times New Roman', Times, serif;
  line-height: 1.8;
  color: #333;
  margin: 10px;
  padding: 10px;
  background-color: white;
}

/* Floating Navigation Bar */
.navbar {
  background-color: rgba(0, 0, 0, 0.3) !important;
  border: none !important;
  position: fixed !important;
  top: 0;
  width: 100%;
  z-index: 1000;
  backdrop-filter: blur(10px);
  -webkit-backdrop-filter: blur(10px);
}

.navbar-default .navbar-brand,
.navbar-default .navbar-nav > li > a {
  color: white !important;
  text-shadow: 1px 1px 2px rgba(0,0,0,0.5);
}

.navbar-default .navbar-brand:hover,
.navbar-default .navbar-nav > li > a:hover {
  color: #f0f0f0 !important;
}

.navbar-default .navbar-toggle {
  border-color: rgba(255,255,255,0.5);
}

.navbar-default .navbar-toggle .icon-bar {
  background-color: white;
}

.navbar-default .navbar-toggle:hover,
.navbar-default .navbar-toggle:focus {
  background-color: rgba(255,255,255,0.1);
}

body {
  padding-top: 70px;
}

/* Content Sections */
.content-section {
  padding: 5px 10px;
  max-width: 1200px;
  margin: 0 auto 10px auto;
  background: rgba(255, 255, 255, 0.7);
  border-radius: 10px;
  box-shadow: 0 4px 15px rgba(0,0,0,0.1);
}

.content-section h1 {
  font-size: 1.8em;
  color: #d97706;
  margin-top: 5px !important;
  margin-bottom: 5px;
  padding-top: 5px !important;
  font-weight: 900;
  font-family: Georgia, 'Times New Roman', Times, serif;
}

.content-section h2 {
  font-size: 1.5em;
  color: #ea580c;
  margin-top: 5px !important;
  margin-bottom: 5px;
  padding-top: 5px !important;
  font-weight: 800;
  font-family: Georgia, 'Times New Roman', Times, serif;
}

.content-section h3 {
  font-size: 1.2em;
  color: #d97706;
  margin-top: 5px !important;
  margin-bottom: 5px;
  padding-top: 5px !important;
  font-weight: 600;
  font-family: Georgia, 'Times New Roman', Times, serif;
}

.content-section p {
  font-size: 1em;
  color: #555;
  margin-bottom: 15px;
  line-height: 1.8;
}

.content-section strong {
  color: #d97706;
  font-weight: 700;
}

.content-section ul, .content-section ol {
  color: #555;
  line-height: 1.8;
  margin-bottom: 15px;
}

.content-section li {
  margin-bottom: 8px;
}

/* Title styling */
h1.title {
  font-family: Georgia, 'Times New Roman', Times, serif !important;
}

/* Code blocks */
pre {
  background: #f5f5f5;
  border-left: 4px solid #d97706;
  padding: 15px;
  border-radius: 5px;
  overflow-x: auto;
}

code {
  color: #d97706;
  background: rgba(217, 119, 6, 0.1);
  padding: 2px 6px;
  border-radius: 3px;
}

/* Tables */
table {
  border-collapse: collapse;
  width: 100%;
  margin: 20px 0;
}

table th {
  background: #d97706;
  color: white;
  padding: 12px;
  text-align: left;
  font-weight: 600;
}

table td {
  padding: 10px;
  border-bottom: 1px solid #e0e0e0;
}

table tr:hover {
  background: rgba(217, 119, 6, 0.05);
}

/* Center all figures */
.figure {
  text-align: center !important;
  margin: 20px auto !important;
}

img {
  display: block !important;
  margin-left: auto !important;
  margin-right: auto !important;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 10,
  fig.asp = 0.6,
  out.width = "90%",
  fig.align = "center"
)

library(tidyverse)
library(ggplot2)
library(knitr)
library(kableExtra)
library(janitor)
library(stringr)
library(maps)
library(patchwork)

# Set serif font for all ggplot text
theme_set(theme_minimal(base_family = "serif") + 
          theme(legend.position = "bottom",
                text = element_text(family = "serif"),
                plot.title = element_text(family = "serif", face = "bold", color = "#d97706"),
                axis.title = element_text(family = "serif"),
                axis.text = element_text(family = "serif"),
                legend.text = element_text(family = "serif"),
                legend.title = element_text(family = "serif")))
```

<div class="content-section">

# Introduction

## Motivation

Alzheimer's disease (AD) is one of the leading causes of death among older adults in the United States and a growing public health challenge as the population ages. According to the Alzheimer's Association, approximately **7.2 million Americans aged 65 or above are projected to be living with Alzheimer's in 2025**, a number expected to rise significantly in coming decades. Understanding how Alzheimer's mortality and self-reported cognitive decline vary across states—and how these patterns relate to social, environmental, and health factors—can inform prevention and policy efforts.

This report explores **temporal and geographic trends in Alzheimer's disease-related cognitive decline prevalence** from 2015 to 2022 and examines how these are associated with contextual factors such as income, education level, and air quality.

## Research Questions

1. How does Alzheimer's-related cognitive prevalence vary across U.S. states and over time?
2. Are there demographic differences (age, sex) in cognitive decline prevalence?
3. What is the association between state-level socioeconomic factors (income, education) and cognitive burden?
4. Is air pollution (PM2.5) associated with cognitive decline at the ecological level?

</div>

<div class="content-section">

# Data and Methods

## Data Sources

```{r data-loading}
# Load Alzheimer's data
raw_az_data <- read.csv("data/2015-2022 Alzheimer Data.csv") |>
  janitor::clean_names()

raw_az_data <- raw_az_data[-nrow(raw_az_data), ]

az_clean <- raw_az_data |>
  filter(
    !is.na(data_value),
    !is.na(year_start),
    !is.na(location_abbr)
  )

# Filter for cognitive decline/impairment
az_cog <- az_clean |>
  filter(
    str_detect(topic, regex("cognitive decline|cognitive impairment",
                            ignore_case = TRUE))
  )

# Create base summary with age and sex groups
az_cog_summary_base <- az_cog |>
  mutate(
    age_group = case_when(
      stratification_category1 == "Age Group" ~ stratification1,
      TRUE ~ "Overall"
    ),
    sex_group = case_when(
      stratification_category2 == "Sex" ~ stratification2,
      TRUE ~ "Overall"
    )
  )

# Filter for 65+ overall
az_cog_65_overall <- az_cog_summary_base |>
  filter(
    age_group == "65 years or older",
    sex_group == "Overall"
  )

# State-year level data
az_state_year <- az_cog_65_overall |>
  group_by(
    year = year_start,
    state_abbr = location_abbr,
    state_name = location_desc
  ) |>
  summarise(
    mean_prev = mean(data_value, na.rm = TRUE),
    .groups = "drop"
  ) |>
  filter(!state_abbr %in% c("US", "GU", "PR", "VI", "DC", "NRE", "SOU", "MDW", "WST"))

# Load socioeconomic data
inc15 <- read_csv("data/median income-2015.csv", skip = 1) |>
  janitor::clean_names() |>
  transmute(
    year = 2015,
    state_name = geographic_area_name,
    median_income = as.numeric(median_income_dollars_estimate_households)
  )

inc16 <- read_csv("data/median income-2016.csv", skip = 1) |>
  janitor::clean_names() |>
  transmute(
    year = 2016,
    state_name = geographic_area_name,
    median_income = as.numeric(median_income_dollars_estimate_households)
  )

income_state <- bind_rows(inc15, inc16) |>
  mutate(state = str_squish(str_extract(state_name, "[^,]+$"))) |>
  group_by(year, state) |>
  summarise(median_state = median(median_income, na.rm = TRUE), .groups = "drop")

# Education data
edu15_raw <- read_csv("data/education2015.csv", skip = 1) |> janitor::clean_names()
edu16_raw <- read_csv("data/education2016.csv", skip = 1) |> janitor::clean_names()

edu_state <- bind_rows(
  edu15_raw |> filter(!str_detect(geographic_area_name, ",")) |>
    transmute(year = 2015, state = geographic_area_name,
              educ_bach = as.numeric(percent_estimate_percent_bachelors_degree_or_higher)),
  edu16_raw |> filter(!str_detect(geographic_area_name, ",")) |>
    transmute(year = 2016, state = geographic_area_name,
              educ_bach = as.numeric(percent_estimate_percent_bachelors_degree_or_higher))
)

# Air quality data
air15 <- read_csv("data/clean_air_state_2015.csv")
air16 <- read_csv("data/clean_air_state_2016.csv")

air_state <- bind_rows(air15, air16) |>
  group_by(year, state_desc) |>
  summarise(pm25_state = mean(pm25_wtd_mean, na.rm = TRUE), .groups = "drop")

# Merge datasets for 2015-2016 analysis
az_cog_65_state <- az_cog_65_overall |>
  filter(year_start %in% c(2015, 2016)) |>
  group_by(year = year_start, state_name = location_desc) |>
  summarise(mean_prev = mean(data_value, na.rm = TRUE), .groups = "drop")

az_income <- az_cog_65_state |>
  left_join(income_state, by = c("year", "state_name" = "state"))

az_edu <- az_cog_65_state |>
  left_join(edu_state, by = c("year", "state_name" = "state"))

az_air <- az_cog_65_state |>
  left_join(air_state, by = c("year", "state_name" = "state_desc"))
```

We utilized data from multiple sources:

| Source | Variables | Years |
|--------|-----------|-------|
| **CDC Alzheimer's Disease and Healthy Aging Data Portal** | Cognitive decline prevalence by state, age, and sex | 2015–2022 |
| **U.S. Census American Community Survey (ACS)** | Median household income, % with bachelor's degree | 2015–2016 |
| **EPA Air Quality System (AQS)** | Annual PM2.5 weighted mean concentration | 2015–2016 |

## Study Population

The primary outcome was the **percentage of adults reporting cognitive decline or impairment**, stratified by:

- **Age groups**: 50–64 years and 65+ years
- **Sex**: Male and Female
- **Geography**: 50 U.S. states (excluding territories and DC)

## Statistical Methods

1. **Descriptive Analysis**: Calculated means, standard deviations, and ranges for prevalence by year, state, age, and sex.

2. **Geographic Visualization**: Created choropleth maps to visualize state-level prevalence patterns.

3. **Regression Analysis**: Used ordinary least squares (OLS) linear regression to examine associations between state-level cognitive prevalence and socioeconomic/environmental factors:

$$\text{Prevalence}_i = \beta_0 + \beta_1 \times \text{Factor}_i + \beta_2 \times \text{Year}_i + \varepsilon_i$$

where Factor represents median income, education (% with bachelor's degree), or PM2.5 concentration.

</div>

<div class="content-section">

# Results

## Temporal Trends

```{r prevalence-year, fig.width=10, fig.asp=0.5}
az_nat_65_overall <- az_cog_65_overall |>
  group_by(year = year_start) |>
  summarise(
    mean_prev = mean(data_value, na.rm = TRUE),
    sd_prev = sd(data_value, na.rm = TRUE),
    .groups = "drop"
  )

ggplot(az_nat_65_overall, aes(x = year, y = mean_prev)) +
  geom_ribbon(aes(ymin = mean_prev - sd_prev, ymax = mean_prev + sd_prev),
              fill = "#d97706", alpha = 0.2) +
  geom_line(color = "#d97706", linewidth = 1.5) +
  geom_point(color = "#d97706", size = 4) +
  geom_text(aes(label = paste0(round(mean_prev, 1), "%")),
            vjust = -1.5, family = "serif", fontface = "bold", color = "#d97706") +
  scale_x_continuous(breaks = seq(2015, 2022, 1)) +
  scale_y_continuous(limits = c(20, 35)) +
  labs(
    title = "National Cognitive Prevalence Among Adults 65+ (2015–2016)",
    x = "Year",
    y = "Mean Prevalence (%)",
    caption = "Shaded area represents ±1 standard deviation across states"
  )
```

The national average prevalence of Alzheimer's-related cognitive indicators among adults aged 65+ remained relatively **stable between 25–27%** from 2015 to 2016. Year-to-year fluctuations were minimal, and no clear upward or downward trend was observed during this period.

## Demographic Differences

### By Age Group

```{r prevalence-age, fig.width=8, fig.asp=0.6}
az_age_desc <- az_cog_summary_base |>
  filter(age_group != "Overall") |>
  group_by(age_group) |>
  summarise(
    mean_value = mean(data_value, na.rm = TRUE),
    sd_value = sd(data_value, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(age_group = factor(age_group, levels = c("50-64 years", "65 years or older")))

ggplot(az_age_desc, aes(x = age_group, y = mean_value, fill = age_group)) +
  geom_col(width = 0.6, alpha = 0.85, color = "black", linewidth = 0.5) +
  geom_errorbar(aes(ymin = mean_value - sd_value, ymax = mean_value + sd_value),
                width = 0.2, color = "#333", linewidth = 0.8) +
  geom_text(aes(label = paste0(round(mean_value, 1), "%")),
            vjust = -0.5, family = "serif", fontface = "bold", size = 5) +
  scale_fill_manual(values = c("50-64 years" = "#d97706", "65 years or older" = "#ea580c")) +
  scale_y_continuous(limits = c(0, 45), expand = c(0, 0)) +
  labs(
    title = "Cognitive Prevalence by Age Group",
    x = "Age Group",
    y = "Mean Prevalence (%)"
  ) +
  theme(legend.position = "none")
```

Adults aged **50–64 years showed higher mean prevalence (~32%)** compared to those 65+ (~26%). This may reflect differences in survey response patterns or the composition of cognitive indicators across age groups.

### By Sex

```{r prevalence-sex, fig.width=8, fig.asp=0.6}
az_sex_desc <- az_cog_summary_base |>
  filter(sex_group != "Overall") |>
  group_by(sex_group) |>
  summarise(
    mean_value = mean(data_value, na.rm = TRUE),
    sd_value = sd(data_value, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(sex_group = factor(sex_group, levels = c("Female", "Male")))

ggplot(az_sex_desc, aes(x = sex_group, y = mean_value, fill = sex_group)) +
  geom_col(width = 0.6, alpha = 0.85, color = "black", linewidth = 0.5) +
  geom_errorbar(aes(ymin = mean_value - sd_value, ymax = mean_value + sd_value),
                width = 0.2, color = "#333", linewidth = 0.8) +
  geom_text(aes(label = paste0(round(mean_value, 1), "%")),
            vjust = -0.5, family = "serif", fontface = "bold", size = 5) +
  scale_fill_manual(values = c("Female" = "#d97706", "Male" = "#2563eb")) +
  scale_y_continuous(limits = c(0, 45), expand = c(0, 0)) +
  labs(
    title = "Cognitive Prevalence by Sex",
    x = "Sex",
    y = "Mean Prevalence (%)"
  ) +
  theme(legend.position = "none")
```

Women reported **higher cognitive burden than men** (approximately 31% vs 27%), consistent with epidemiological literature showing higher Alzheimer's prevalence among females.

## Geographic Patterns

```{r maps, fig.width=14, fig.asp=0.5}
us_map <- map_data("state") |> as_tibble()

az_2015 <- az_state_year |> 
  filter(year == 2015) |>
  mutate(region = str_to_lower(state_name)) |>
  inner_join(us_map, by = "region")

az_2016 <- az_state_year |> 
  filter(year == 2016) |>
  mutate(region = str_to_lower(state_name)) |>
  inner_join(us_map, by = "region")

p1 <- ggplot(az_2015, aes(long, lat, group = group, fill = mean_prev)) +
  geom_polygon(color = "black", linewidth = 0.5) +
  geom_path(data = us_map, aes(long, lat, group = group), 
            color = "black", linewidth = 0.8, inherit.aes = FALSE) +
  coord_fixed(1.3) +
  scale_fill_viridis_c(option = "C", limits = c(18, 35)) +
  labs(title = "2015", fill = "Prevalence (%)") +
  theme_void(base_family = "serif") +
  theme(
    plot.title = element_text(size = 16, face = "bold", color = "#d97706", hjust = 0.5),
    legend.position = "right"
  )

p2 <- ggplot(az_2016, aes(long, lat, group = group, fill = mean_prev)) +
  geom_polygon(color = "black", linewidth = 0.5) +
  geom_path(data = us_map, aes(long, lat, group = group), 
            color = "black", linewidth = 0.8, inherit.aes = FALSE) +
  coord_fixed(1.3) +
  scale_fill_viridis_c(option = "C", limits = c(18, 35)) +
  labs(title = "2016", fill = "Prevalence (%)") +
  theme_void(base_family = "serif") +
  theme(
    plot.title = element_text(size = 16, face = "bold", color = "#d97706", hjust = 0.5),
    legend.position = "right"
  )

p1 + p2 + plot_annotation(
  title = "State-Level Cognitive Prevalence Among Adults 65+",
  theme = theme(
    plot.title = element_text(family = "serif", size = 18, face = "bold", 
                              color = "#d97706", hjust = 0.5)
  )
)
```

Geographic patterns revealed **regional disparities** in cognitive prevalence:

- **Higher prevalence**: Southern and Southeastern states (e.g., Mississippi, Louisiana, Alabama, Georgia)
- **Lower prevalence**: Western states (e.g., Utah, Colorado) and Northeastern states (e.g., Massachusetts, Connecticut)

These patterns remained consistent between 2015 and 2016.

## Socioeconomic Factor Analysis

### Income

```{r income-plot, fig.width=10, fig.asp=0.6}
ggplot(az_income |> filter(!is.na(median_state)),
       aes(x = median_state, y = mean_prev, color = factor(year))) +
  geom_point(alpha = 0.7, size = 3) +
  geom_smooth(method = "lm", se = TRUE, alpha = 0.2) +
  scale_color_manual(values = c("#d97706", "#2563eb")) +
  scale_x_continuous(labels = scales::dollar_format()) +
  labs(
    title = "State-Level Cognitive Burden vs Median Income (2015–2016)",
    x = "State Median Household Income",
    y = "Cognitive Prevalence (%)",
    color = "Year"
  )
```

```{r income-model, echo=FALSE}
model_inc <- lm(mean_prev ~ median_state + factor(year), data = az_income)
inc_coef <- round(coef(model_inc)[2] * 10000, 3)
inc_r2 <- round(summary(model_inc)$r.squared, 3)
```

Higher state-level median income was associated with **lower cognitive prevalence**. The regression coefficient was **`r inc_coef` percentage points per $10,000** increase in income. However, the model R² was only **`r inc_r2`**, indicating income explains only a small fraction of between-state variation.

### Education

```{r education-plot, fig.width=10, fig.asp=0.6}
ggplot(az_edu |> filter(!is.na(educ_bach)),
       aes(x = educ_bach, y = mean_prev, color = factor(year))) +
  geom_point(alpha = 0.7, size = 3) +
  geom_smooth(method = "lm", se = TRUE, alpha = 0.2) +
  scale_color_manual(values = c("#d97706", "#2563eb")) +
  labs(
    title = "State-Level Cognitive Burden vs Education (2015–2016)",
    x = "% Adults with Bachelor's Degree or Higher",
    y = "Cognitive Prevalence (%)",
    color = "Year"
  )
```

```{r education-model, echo=FALSE}
model_edu <- lm(mean_prev ~ educ_bach + factor(year), data = az_edu)
edu_coef <- round(coef(model_edu)[2], 3)
edu_r2 <- round(summary(model_edu)$r.squared, 3)
```

States with higher educational attainment showed **modestly lower cognitive prevalence**. A 1-percentage-point increase in the share of adults with a bachelor's degree was associated with approximately **`r edu_coef` percentage points lower** prevalence. The model R² was **`r edu_r2`**.

### Air Quality (PM2.5)

```{r air-plot, fig.width=10, fig.asp=0.6}
ggplot(az_air |> filter(!is.na(pm25_state)),
       aes(x = pm25_state, y = mean_prev, color = factor(year))) +
  geom_point(alpha = 0.7, size = 3) +
  geom_smooth(method = "lm", se = TRUE, alpha = 0.2) +
  scale_color_manual(values = c("#d97706", "#2563eb")) +
  labs(
    title = "State-Level Cognitive Burden vs PM2.5 (2015–2016)",
    x = "PM2.5 Weighted Mean (µg/m³)",
    y = "Cognitive Prevalence (%)",
    color = "Year"
  )
```

```{r air-model, echo=FALSE}
model_air <- lm(mean_prev ~ pm25_state + factor(year), data = az_air)
air_coef <- round(coef(model_air)[2], 3)
air_pval <- round(summary(model_air)$coefficients[2, 4], 3)
air_r2 <- round(summary(model_air)$r.squared, 3)
```

PM2.5 concentration showed **no statistically significant association** with cognitive prevalence at the state level (β = `r air_coef`, p = `r air_pval`). The model R² was only **`r air_r2`**.

## Summary of Regression Results

```{r regression-summary, echo=FALSE}
summary_table <- data.frame(
  Factor = c("Median Income (per $10,000)", "Education (% BA+)", "PM2.5 (µg/m³)"),
  Coefficient = c(
    round(coef(model_inc)[2] * 10000, 3),
    round(coef(model_edu)[2], 3),
    round(coef(model_air)[2], 3)
  ),
  `R-squared` = c(
    round(summary(model_inc)$r.squared, 3),
    round(summary(model_edu)$r.squared, 3),
    round(summary(model_air)$r.squared, 3)
  ),
  Interpretation = c(
    "Higher income → lower prevalence",
    "Higher education → lower prevalence",
    "No significant association"
  )
)

kable(summary_table, col.names = c("Factor", "β Coefficient", "R²", "Interpretation"),
      align = c("l", "r", "r", "l"),
      caption = "Summary of State-Level Regression Models (2015–2016)") |>
  kable_styling(bootstrap_options = c("striped", "hover"),
                full_width = FALSE,
                position = "center") |>
  column_spec(1, bold = TRUE)
```

</div>

<div class="content-section">

# Discussion

## Key Findings

1. **Temporal Stability**: National cognitive prevalence among adults 65+ remained stable at approximately 25–27% from 2015–2016, with no clear trend.

2. **Demographic Disparities**: Women consistently reported higher cognitive burden than men. The 50–64 age group showed higher prevalence than those 65+.

3. **Geographic Variation**: Substantial state-level variation exists, with Southern/Southeastern states showing higher prevalence and Western/Northeastern states showing lower prevalence.

4. **Weak Socioeconomic Associations**: While higher income and education were associated with lower cognitive prevalence, these factors explained only a small fraction (<5%) of between-state variation.

5. **No Air Quality Association**: PM2.5 was not significantly associated with cognitive prevalence at the ecological level.

## Limitations

1. **Ecological Fallacy**: This analysis uses state-level aggregates, so associations observed at the state level may not hold at the individual level.

2. **Self-Reported Data**: Cognitive decline prevalence is based on self-report (BRFSS), which may be subject to reporting bias.

3. **Limited Covariates**: We examined only a few socioeconomic and environmental factors; other determinants (healthcare access, genetics, comorbidities) were not included.

4. **Short Time Period for Factor Analysis**: Socioeconomic and air quality data were only available for 2015–2016, limiting temporal analysis.

5. **Measurement Error**: Both predictor and outcome variables are measured with error, potentially attenuating associations.

## Public Health Implications

- Geographic disparities suggest targeted interventions may be needed in high-prevalence regions
- The weak associations with socioeconomic factors highlight that multiple determinants beyond income and education influence cognitive health
- Continued surveillance is essential to detect emerging trends

</div>

<div class="content-section">

# Conclusions

This analysis reveals significant **geographic disparities** in Alzheimer's-related cognitive prevalence across U.S. states, with Southern regions consistently showing higher burden. While state-level income and education showed expected negative associations with cognitive prevalence, these factors explained only a small portion of the variation, suggesting that other unmeasured factors play important roles.

Future research should:

- Incorporate individual-level data to avoid ecological bias
- Examine additional risk factors (cardiovascular disease, diabetes, social isolation)
- Explore longitudinal changes with more years of socioeconomic data
- Investigate why certain geographic regions show persistently higher prevalence

</div>

<div class="content-section">

# References

1. Alzheimer's Association. (2024). 2024 Alzheimer's Disease Facts and Figures. *Alzheimer's & Dementia*, 20(5).

2. CDC. Alzheimer's Disease and Healthy Aging Data Portal. https://data.cdc.gov/Healthy-Aging/Alzheimer-s-Disease-and-Healthy-Aging-Data/hfr9-rurv

3. U.S. Census Bureau. American Community Survey. https://data.census.gov

4. U.S. Environmental Protection Agency. Air Quality System Data. https://www.epa.gov/outdoor-air-quality-data

5. Livingston, G., et al. (2020). Dementia prevention, intervention, and care: 2020 report of the Lancet Commission. *The Lancet*, 396(10248), 413-446.

</div>