---
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    theme: flatly
---

<style>
/* Global Styles */
html, body {
  font-family: Georgia, 'Times New Roman', Times, serif;
  line-height: 1.8;
  color: #333;
  margin: 10px;
  padding: 10px;
  background-color: white;
}

/* Floating Navigation Bar */
.navbar {
  background-color: rgba(0, 0, 0, 0.3) !important;
  border: none !important;
  position: fixed !important;
  top: 0;
  width: 100%;
  z-index: 1000;
  backdrop-filter: blur(10px);
  -webkit-backdrop-filter: blur(10px);
}

.navbar-default .navbar-brand,
.navbar-default .navbar-nav > li > a {
  color: white !important;
  text-shadow: 1px 1px 2px rgba(0,0,0,0.5);
}

.navbar-default .navbar-brand:hover,
.navbar-default .navbar-nav > li > a:hover {
  color: #f0f0f0 !important;
}

.navbar-default .navbar-toggle {
  border-color: rgba(255,255,255,0.5);
}

.navbar-default .navbar-toggle .icon-bar {
  background-color: white;
}

.navbar-default .navbar-toggle:hover,
.navbar-default .navbar-toggle:focus {
  background-color: rgba(255,255,255,0.1);
}

body {
  padding-top: 70px;
}

/* Content Sections */
.content-section {
  padding: 5px 10px;
  max-width: 1200px;
  margin: 0 auto 10px auto;
  background: rgba(255, 255, 255, 0.7);
  border-radius: 10px;
  box-shadow: 0 4px 15px rgba(0,0,0,0.1);
}

.content-section h1 {
  font-size: 1.8em;
  color: #d97706;
  margin-top: 5px !important;
  margin-bottom: 5px;
  padding-top: 5px !important;
  font-weight: 900;
  font-family: Georgia, 'Times New Roman', Times, serif;
}

.content-section h2 {
  font-size: 1.5em;
  color: #ea580c;
  margin-top: 5px !important;
  margin-bottom: 5px;
  padding-top: 5px !important;
  font-weight: 800;
  font-family: Georgia, 'Times New Roman', Times, serif;
}

.content-section h3 {
  font-size: 1.2em;
  color: #d97706;
  margin-top: 5px !important;
  margin-bottom: 5px;
  padding-top: 5px !important;
  font-weight: 600;
  font-family: Georgia, 'Times New Roman', Times, serif;
}

.content-section p {
  font-size: 1em;
  color: #555;
  margin-bottom: 15px;
  line-height: 1.8;
}

.content-section strong {
  color: #d97706;
  font-weight: 700;
}

.content-section ul, .content-section ol {
  color: #555;
  line-height: 1.8;
  margin-bottom: 15px;
}

.content-section li {
  margin-bottom: 8px;
}

/* Shiny app iframe styling */
.shiny-frame {
  border: 2px solid #d97706;
  border-radius: 10px;
  box-shadow: 0 4px 15px rgba(217, 119, 6, 0.2);
  margin: 20px auto;
  display: block;
}

/* Interactive visualization containers */
.viz-container {
  background: linear-gradient(135deg, #fff9f0 0%, #ffffff 100%);
  border-radius: 10px;
  padding: 15px;
  margin: 15px 0;
  border-left: 4px solid #d97706;
}

/* Filter controls styling */
.filter-controls {
  background: #f8f4f0;
  padding: 15px;
  border-radius: 8px;
  margin-bottom: 20px;
}

.filter-controls label {
  color: #d97706;
  font-weight: 600;
}

/* Code blocks */
pre {
  background: #f5f5f5;
  border-left: 4px solid #d97706;
  padding: 15px;
  border-radius: 5px;
  overflow-x: auto;
}

code {
  color: #d97706;
  background: rgba(217, 119, 6, 0.1);
  padding: 2px 6px;
  border-radius: 3px;
}

/* Tables */
table {
  border-collapse: collapse;
  width: 100%;
  margin: 20px 0;
}

table th {
  background: #d97706;
  color: white;
  padding: 12px;
  text-align: left;
  font-weight: 600;
}

table td {
  padding: 10px;
  border-bottom: 1px solid #e0e0e0;
}

table tr:hover {
  background: rgba(217, 119, 6, 0.05);
}

/* Center all figures */
.figure {
  text-align: center !important;
  margin: 20px auto !important;
}

img {
  display: block !important;
  margin-left: auto !important;
  margin-right: auto !important;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 10,
  fig.asp = 0.6,
  out.width = "100%"
)

library(tidyverse)
library(plotly)
library(crosstalk)
library(DT)
library(viridis)
library(htmltools)
library(knitr)
library(kableExtra)
library(janitor)
library(stringr)

# Custom plotly theme colors
orange_palette <- c("#d97706", "#ea580c", "#f59e0b", "#fbbf24", "#fcd34d")
```

<div class="content-section">

# Interactive Visualizations

This page features **interactive visualizations** that allow you to explore Alzheimer's disease and cognitive decline data across the United States. Hover over data points for details, zoom in on regions of interest, and filter data dynamically.

**Features include:**

- **Interactive Maps**: Explore state-level cognitive burden with hover tooltips
- **Dynamic Time Series**: Track prevalence trends over time
- **Linked Scatter Plots**: Examine relationships between socioeconomic factors and cognitive outcomes
- **Data Explorer**: Filter and search the underlying data
- **Shiny Dashboard**: A comprehensive interactive tool for in-depth analysis

</div>

<div class="content-section">

# Data Preparation

```{r data-loading}
# Load Alzheimer's data
raw_az_data <- read.csv("data/2015-2022 Alzheimer Data.csv") |>
  janitor::clean_names()

raw_az_data <- raw_az_data[-nrow(raw_az_data), ]

az_clean <- raw_az_data |>
  filter(
    !is.na(data_value),
    !is.na(year_start),
    !is.na(location_abbr)
  )

# Filter for cognitive decline/impairment
az_cog <- az_clean |>
  filter(
    str_detect(topic, regex("cognitive decline|cognitive impairment",
                            ignore_case = TRUE))
  )

# Create base summary with age and sex groups
az_cog_summary_base <- az_cog |>
  mutate(
    age_group = case_when(
      stratification_category1 == "Age Group" ~ stratification1,
      TRUE ~ "Overall"
    ),
    sex_group = case_when(
      stratification_category2 == "Sex" ~ stratification2,
      TRUE ~ "Overall"
    )
  )

# Filter for 65+ overall
az_cog_65_overall <- az_cog_summary_base |>
  filter(
    age_group == "65 years or older",
    sex_group == "Overall"
  )

# State-year level data for all years
az_state_year_all <- az_cog_65_overall |>
  group_by(
    year = year_start,
    state_abbr = location_abbr,
    state_name = location_desc
  ) |>
  summarise(
    mean_prev = mean(data_value, na.rm = TRUE),
    .groups = "drop"
  ) |>
  filter(!state_abbr %in% c("US", "GU", "PR", "VI", "DC", "NRE", "SOU", "MDW", "WST"))

# Load and prepare socioeconomic data for 2015-2016
inc15 <- read_csv("data/median income-2015.csv", skip = 1) |>
  janitor::clean_names() |>
  transmute(
    year = 2015,
    state_name = geographic_area_name,
    median_income = as.numeric(median_income_dollars_estimate_households)
  )

inc16 <- read_csv("data/median income-2016.csv", skip = 1) |>
  janitor::clean_names() |>
  transmute(
    year = 2016,
    state_name = geographic_area_name,
    median_income = as.numeric(median_income_dollars_estimate_households)
  )

income_state <- bind_rows(inc15, inc16) |>
  mutate(state = str_squish(str_extract(state_name, "[^,]+$"))) |>
  group_by(year, state) |>
  summarise(
    median_state = median(median_income, na.rm = TRUE),
    .groups = "drop"
  )

# Education data
edu15_raw <- read_csv("data/education2015.csv", skip = 1) |> janitor::clean_names()
edu16_raw <- read_csv("data/education2016.csv", skip = 1) |> janitor::clean_names()

edu_state <- bind_rows(
  edu15_raw |> filter(!str_detect(geographic_area_name, ",")) |>
    transmute(year = 2015, state = geographic_area_name,
              educ_bach = as.numeric(percent_estimate_percent_bachelors_degree_or_higher)),
  edu16_raw |> filter(!str_detect(geographic_area_name, ",")) |>
    transmute(year = 2016, state = geographic_area_name,
              educ_bach = as.numeric(percent_estimate_percent_bachelors_degree_or_higher))
)

# Air quality data
air15 <- read_csv("data/clean_air_state_2015.csv")
air16 <- read_csv("data/clean_air_state_2016.csv")

air_state <- bind_rows(air15, air16) |>
  group_by(year, state_desc) |>
  summarise(pm25_state = mean(pm25_wtd_mean, na.rm = TRUE), .groups = "drop")

# Combined dataset for 2015-2016
az_combined <- az_state_year_all |>
  filter(year %in% c(2015, 2016)) |>
  left_join(income_state, by = c("year", "state_name" = "state")) |>
  left_join(edu_state, by = c("year", "state_name" = "state")) |>
  left_join(air_state, by = c("year", "state_name" = "state_desc"))
```

</div>

<div class="content-section">

# Interactive Choropleth Maps

Explore Alzheimer's-related cognitive prevalence across U.S. states. **Hover over states** to see detailed information, and use the **dropdown** to switch between years.

```{r interactive-map, fig.height=6}
# Prepare data for plotly choropleth - one map per year with dropdown
years <- sort(unique(az_state_year_all$year))

# Get data for the first year (default view)
first_year_data <- az_state_year_all |> filter(year == years[1])

# Create the base map
fig_map <- plot_ly(
  type = "choropleth",
  locationmode = "USA-states",
  locations = first_year_data$state_abbr,
  z = first_year_data$mean_prev,
  text = paste0("<b>", first_year_data$state_name, "</b><br>",
                "Year: ", first_year_data$year, "<br>",
                "Prevalence: ", round(first_year_data$mean_prev, 1), "%"),
  hoverinfo = "text",
  colorscale = list(
    c(0, "#fef3c7"),
    c(0.25, "#fcd34d"),
    c(0.5, "#f59e0b"),
    c(0.75, "#d97706"),
    c(1, "#92400e")
  ),
  zmin = min(az_state_year_all$mean_prev, na.rm = TRUE),
  zmax = max(az_state_year_all$mean_prev, na.rm = TRUE),
  colorbar = list(
    title = list(text = "Prevalence (%)", font = list(family = "Georgia")),
    tickfont = list(family = "Georgia")
  )
)

# Create dropdown buttons for each year
buttons <- lapply(years, function(yr) {
  year_data <- az_state_year_all |> filter(year == yr)
  list(
    method = "restyle",
    args = list(list(
      z = list(year_data$mean_prev),
      locations = list(year_data$state_abbr),
      text = list(paste0("<b>", year_data$state_name, "</b><br>",
                         "Year: ", year_data$year, "<br>",
                         "Prevalence: ", round(year_data$mean_prev, 1), "%"))
    )),
    label = as.character(yr)
  )
})

fig_map <- fig_map |>
  layout(
    title = list(
      text = "<b>Alzheimer's-Related Cognitive Prevalence by State (Age 65+)</b>",
      font = list(family = "Georgia", size = 18, color = "#d97706"),
      y = 0.98,
      x = 0.5,
      xanchor = "center"
    ),
    geo = list(
      scope = "usa",
      showlakes = TRUE,
      lakecolor = toRGB("white"),
      bgcolor = "rgba(0,0,0,0)"
    ),
    paper_bgcolor = "rgba(0,0,0,0)",
    plot_bgcolor = "rgba(0,0,0,0)",
    font = list(family = "Georgia"),
    margin = list(t = 80),
    updatemenus = list(
      list(
        type = "dropdown",
        active = 0,
        buttons = buttons,
        x = 0.02,
        y = 0.95,
        xanchor = "left",
        yanchor = "top",
        font = list(family = "Georgia"),
        bgcolor = "#fff",
        bordercolor = "#d97706",
        pad = list(r = 10, t = 10)
      )
    ),
    annotations = list(
      list(
        text = "<b>Select Year:</b>",
        x = 0.02,
        y = 1.0,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(family = "Georgia", size = 12, color = "#d97706")
      )
    )
  )

fig_map
```

The interactive map reveals regional patterns in cognitive decline prevalence. **Southern and Southeastern states** consistently show higher prevalence rates, while **Western and Northeastern states** tend to have lower rates. Use the year slider to observe temporal changes.

</div>

<div class="content-section">

# Time Series Analysis

## National Trends Over Time

Track the **national average** cognitive prevalence from 2015 to 2022. Click on legend items to show/hide specific metrics.

```{r time-series, fig.height=5}
# National yearly summary
national_trend <- az_state_year_all |>
  group_by(year) |>
  summarise(
    mean_prev = mean(mean_prev, na.rm = TRUE),
    sd_prev = sd(mean_prev, na.rm = TRUE),
    min_prev = min(mean_prev, na.rm = TRUE),
    max_prev = max(mean_prev, na.rm = TRUE),
    .groups = "drop"
  )

fig_trend <- plot_ly(national_trend, x = ~year) |>
  add_ribbons(
    ymin = ~mean_prev - sd_prev,
    ymax = ~mean_prev + sd_prev,
    fillcolor = "rgba(217, 119, 6, 0.2)",
    line = list(color = "transparent"),
    name = "±1 SD",
    hoverinfo = "skip"
  ) |>
  add_trace(
    y = ~mean_prev,
    type = "scatter",
    mode = "lines+markers",
    line = list(color = "#d97706", width = 3),
    marker = list(color = "#d97706", size = 10, 
                  line = list(color = "white", width = 2)),
    name = "National Mean",
    hovertemplate = paste0(
      "<b>Year: %{x}</b><br>",
      "Mean Prevalence: %{y:.1f}%<br>",
      "<extra></extra>"
    )
  ) |>
  add_trace(
    y = ~min_prev,
    type = "scatter",
    mode = "lines",
    line = list(color = "#fbbf24", width = 1, dash = "dot"),
    name = "State Min",
    hovertemplate = "Min: %{y:.1f}%<extra></extra>"
  ) |>
  add_trace(
    y = ~max_prev,
    type = "scatter",
    mode = "lines",
    line = list(color = "#92400e", width = 1, dash = "dot"),
    name = "State Max",
    hovertemplate = "Max: %{y:.1f}%<extra></extra>"
  ) |>
  layout(
    title = list(
      text = "<b>National Cognitive Prevalence Trends (2015-2022)</b>",
      font = list(family = "Georgia", size = 16, color = "#d97706")
    ),
    xaxis = list(
      title = "Year",
      tickfont = list(family = "Georgia"),
      titlefont = list(family = "Georgia", size = 14),
      dtick = 1
    ),
    yaxis = list(
      title = "Prevalence (%)",
      tickfont = list(family = "Georgia"),
      titlefont = list(family = "Georgia", size = 14),
      range = c(15, 35)
    ),
    legend = list(
      orientation = "h",
      x = 0.5,
      xanchor = "center",
      y = -0.15,
      font = list(family = "Georgia")
    ),
    paper_bgcolor = "rgba(0,0,0,0)",
    plot_bgcolor = "rgba(0,0,0,0)",
    font = list(family = "Georgia"),
    hovermode = "x unified"
  )

fig_trend
```

## State Comparison Over Time

Select multiple states to compare their cognitive prevalence trajectories.

```{r state-comparison, fig.height=6}
# Top and bottom 5 states by average prevalence
state_avg <- az_state_year_all |>
  group_by(state_name, state_abbr) |>
  summarise(avg_prev = mean(mean_prev, na.rm = TRUE), .groups = "drop") |>
  arrange(desc(avg_prev))

highlight_states <- c(
  head(state_avg$state_abbr, 5),  # Top 5
  tail(state_avg$state_abbr, 5)   # Bottom 5
)

state_trends <- az_state_year_all |>
  filter(state_abbr %in% highlight_states) |>
  mutate(
    category = case_when(
      state_abbr %in% head(state_avg$state_abbr, 5) ~ "Highest Prevalence",
      TRUE ~ "Lowest Prevalence"
    )
  )

fig_states <- plot_ly(state_trends, x = ~year, y = ~mean_prev, 
                      color = ~state_name, colors = viridis(10)) |>
  add_trace(
    type = "scatter",
    mode = "lines+markers",
    marker = list(size = 8),
    line = list(width = 2),
    hovertemplate = paste0(
      "<b>%{text}</b><br>",
      "Year: %{x}<br>",
      "Prevalence: %{y:.1f}%<br>",
      "<extra></extra>"
    ),
    text = ~state_name
  ) |>
  layout(
    title = list(
      text = "<b>State Comparison: Highest vs Lowest Prevalence States</b>",
      font = list(family = "Georgia", size = 16, color = "#d97706")
    ),
    xaxis = list(
      title = "Year",
      tickfont = list(family = "Georgia"),
      titlefont = list(family = "Georgia"),
      dtick = 1
    ),
    yaxis = list(
      title = "Prevalence (%)",
      tickfont = list(family = "Georgia"),
      titlefont = list(family = "Georgia")
    ),
    legend = list(
      font = list(family = "Georgia"),
      orientation = "v",
      x = 1.02,
      y = 0.5
    ),
    paper_bgcolor = "rgba(0,0,0,0)",
    plot_bgcolor = "rgba(0,0,0,0)",
    font = list(family = "Georgia"),
    hovermode = "closest"
  )

fig_states
```

</div>

<div class="content-section">

# Socioeconomic Factor Analysis

## Interactive Scatter Plots with Linked Brushing

The following visualizations use **linked brushing** – select points in one plot and see them highlighted across all plots. This helps identify states that may be outliers across multiple dimensions.

```{r linked-scatter, fig.height=8}
# Create shared data for crosstalk
shared_data <- SharedData$new(
  az_combined |> 
    filter(!is.na(median_state) & !is.na(educ_bach) & !is.na(pm25_state)) |>
    mutate(
      year_factor = factor(year),
      tooltip = paste0(
        "<b>", state_name, " (", year, ")</b><br>",
        "Prevalence: ", round(mean_prev, 1), "%<br>",
        "Income: $", format(round(median_state), big.mark = ","), "<br>",
        "Education: ", round(educ_bach, 1), "%<br>",
        "PM2.5: ", round(pm25_state, 1), " µg/m³"
      )
    ),
  key = ~paste(state_abbr, year)
)

# Income vs Prevalence
p1 <- plot_ly(shared_data, x = ~median_state, y = ~mean_prev, 
              color = ~year_factor, colors = c("#d97706", "#2563eb")) |>
  add_markers(
    marker = list(size = 12, opacity = 0.7,
                  line = list(color = "white", width = 1)),
    hovertemplate = "%{text}<extra></extra>",
    text = ~tooltip
  ) |>
  layout(
    xaxis = list(title = "Median Household Income ($)", tickfont = list(family = "Georgia")),
    yaxis = list(title = "Prevalence (%)", tickfont = list(family = "Georgia")),
    showlegend = FALSE
  )

# Education vs Prevalence
p2 <- plot_ly(shared_data, x = ~educ_bach, y = ~mean_prev,
              color = ~year_factor, colors = c("#d97706", "#2563eb")) |>
  add_markers(
    marker = list(size = 12, opacity = 0.7,
                  line = list(color = "white", width = 1)),
    hovertemplate = "%{text}<extra></extra>",
    text = ~tooltip
  ) |>
  layout(
    xaxis = list(title = "% with Bachelor's Degree", tickfont = list(family = "Georgia")),
    yaxis = list(title = "Prevalence (%)", tickfont = list(family = "Georgia")),
    showlegend = FALSE
  )

# PM2.5 vs Prevalence
p3 <- plot_ly(shared_data, x = ~pm25_state, y = ~mean_prev,
              color = ~year_factor, colors = c("#d97706", "#2563eb")) |>
  add_markers(
    marker = list(size = 12, opacity = 0.7,
                  line = list(color = "white", width = 1)),
    hovertemplate = "%{text}<extra></extra>",
    text = ~tooltip
  ) |>
  layout(
    xaxis = list(title = "PM2.5 (µg/m³)", tickfont = list(family = "Georgia")),
    yaxis = list(title = "Prevalence (%)", tickfont = list(family = "Georgia")),
    legend = list(
      title = list(text = "Year"),
      font = list(family = "Georgia")
    )
  )

# Combine plots
subplot(p1, p2, p3, nrows = 1, shareY = TRUE, titleX = TRUE) |>
  layout(
    title = list(
      text = "<b>Cognitive Prevalence vs Socioeconomic & Environmental Factors</b>",
      font = list(family = "Georgia", size = 16, color = "#d97706"),
      y = 0.98
    ),
    annotations = list(
      list(x = 0.12, y = 1.08, text = "<b>Income Effect</b>", 
           showarrow = FALSE, xref = "paper", yref = "paper",
           font = list(family = "Georgia", size = 13, color = "#ea580c")),
      list(x = 0.5, y = 1.08, text = "<b>Education Effect</b>",
           showarrow = FALSE, xref = "paper", yref = "paper",
           font = list(family = "Georgia", size = 13, color = "#ea580c")),
      list(x = 0.88, y = 1.08, text = "<b>Air Quality Effect</b>",
           showarrow = FALSE, xref = "paper", yref = "paper",
           font = list(family = "Georgia", size = 13, color = "#ea580c"))
    ),
    paper_bgcolor = "rgba(0,0,0,0)",
    plot_bgcolor = "rgba(0,0,0,0)",
    font = list(family = "Georgia"),
    hovermode = "closest"
  ) |>
  highlight(on = "plotly_click", off = "plotly_doubleclick", 
            color = "#ef4444", opacityDim = 0.3)
```

**Key Observations:**

- **Income**: States with higher median household income tend to show slightly lower cognitive prevalence
- **Education**: Higher educational attainment is associated with modestly lower prevalence rates  
- **Air Quality**: PM2.5 levels show minimal correlation with cognitive outcomes at the state level

</div>

<div class="content-section">

# Distribution Analysis

## Prevalence by Demographics

```{r demographic-plots, fig.height=5}
# Prepare demographic summaries
az_by_age <- az_cog_summary_base |>
  filter(age_group != "Overall") |>
  group_by(year = year_start, age_group) |>
  summarise(mean_value = mean(data_value, na.rm = TRUE), .groups = "drop")

az_by_sex <- az_cog_summary_base |>
  filter(sex_group != "Overall") |>
  group_by(year = year_start, sex_group) |>
  summarise(mean_value = mean(data_value, na.rm = TRUE), .groups = "drop")

# Age group plot
p_age <- plot_ly(az_by_age, x = ~year, y = ~mean_value, 
                 color = ~age_group, colors = c("#d97706", "#2563eb")) |>
  add_trace(
    type = "bar",
    hovertemplate = paste0(
      "<b>%{text}</b><br>",
      "Year: %{x}<br>",
      "Prevalence: %{y:.1f}%<br>",
      "<extra></extra>"
    ),
    text = ~age_group
  ) |>
  layout(
    barmode = "group",
    xaxis = list(title = "Year", tickfont = list(family = "Georgia"), dtick = 1),
    yaxis = list(title = "Prevalence (%)", tickfont = list(family = "Georgia")),
    legend = list(title = list(text = "Age Group"), font = list(family = "Georgia"))
  )

# Sex group plot  
p_sex <- plot_ly(az_by_sex, x = ~year, y = ~mean_value,
                 color = ~sex_group, colors = c("#d97706", "#2563eb")) |>
  add_trace(
    type = "bar",
    hovertemplate = paste0(
      "<b>%{text}</b><br>",
      "Year: %{x}<br>",
      "Prevalence: %{y:.1f}%<br>",
      "<extra></extra>"
    ),
    text = ~sex_group
  ) |>
  layout(
    barmode = "group",
    xaxis = list(title = "Year", tickfont = list(family = "Georgia"), dtick = 1),
    yaxis = list(title = "Prevalence (%)", tickfont = list(family = "Georgia")),
    legend = list(title = list(text = "Sex"), font = list(family = "Georgia"))
  )

subplot(p_age, p_sex, nrows = 1, shareY = TRUE, titleX = TRUE) |>
  layout(
    title = list(
      text = "<b>Prevalence by Age Group and Sex Over Time</b>",
      font = list(family = "Georgia", size = 16, color = "#d97706")
    ),
    paper_bgcolor = "rgba(0,0,0,0)",
    plot_bgcolor = "rgba(0,0,0,0)",
    font = list(family = "Georgia")
  )
```

## State-Level Distribution

```{r boxplot, fig.height=5}
# Create boxplot by year
fig_box <- plot_ly(az_state_year_all, x = ~factor(year), y = ~mean_prev) |>
  add_boxplot(
    fillcolor = "rgba(217, 119, 6, 0.5)",
    line = list(color = "#d97706"),
    marker = list(color = "#d97706", outliercolor = "#92400e"),
    hoverinfo = "y",
    name = ""
  ) |>
  add_trace(
    data = az_state_year_all |> group_by(year) |> summarise(mean_prev = mean(mean_prev)),
    x = ~factor(year),
    y = ~mean_prev,
    type = "scatter",
    mode = "markers",
    marker = list(color = "#ef4444", size = 12, symbol = "diamond"),
    name = "National Mean",
    hovertemplate = "National Mean: %{y:.1f}%<extra></extra>"
  ) |>
  layout(
    title = list(
      text = "<b>Distribution of State-Level Prevalence by Year</b>",
      font = list(family = "Georgia", size = 16, color = "#d97706")
    ),
    xaxis = list(
      title = "Year",
      tickfont = list(family = "Georgia"),
      titlefont = list(family = "Georgia")
    ),
    yaxis = list(
      title = "Prevalence (%)",
      tickfont = list(family = "Georgia"),
      titlefont = list(family = "Georgia")
    ),
    showlegend = TRUE,
    legend = list(
      orientation = "h",
      x = 0.5,
      xanchor = "center",
      y = -0.15,
      font = list(family = "Georgia")
    ),
    paper_bgcolor = "rgba(0,0,0,0)",
    plot_bgcolor = "rgba(0,0,0,0)",
    font = list(family = "Georgia")
  )

fig_box
```

</div>

<div class="content-section">

# Data Explorer

Use the interactive table below to **search, filter, and sort** the underlying data. Click column headers to sort, use the search box to find specific states, and adjust entries per page.

```{r data-table}
# Prepare comprehensive dataset for display
display_data <- az_state_year_all |>
  left_join(income_state, by = c("year", "state_name" = "state")) |>
  left_join(edu_state, by = c("year", "state_name" = "state")) |>
  left_join(air_state, by = c("year", "state_name" = "state_desc")) |>
  select(
    Year = year,
    State = state_name,
    `State Code` = state_abbr,
    `Prevalence (%)` = mean_prev,
    `Median Income ($)` = median_state,
    `Education (% BA+)` = educ_bach,
    `PM2.5 (µg/m³)` = pm25_state
  ) |>
  mutate(
    `Prevalence (%)` = round(`Prevalence (%)`, 1),
    `Median Income ($)` = ifelse(!is.na(`Median Income ($)`), 
                                  format(round(`Median Income ($)`), big.mark = ","), 
                                  NA),
    `Education (% BA+)` = round(`Education (% BA+)`, 1),
    `PM2.5 (µg/m³)` = round(`PM2.5 (µg/m³)`, 1)
  )

datatable(
  display_data,
  filter = "top",
  options = list(
    pageLength = 15,
    autoWidth = TRUE,
    dom = 'Bfrtip',
    scrollX = TRUE,
    columnDefs = list(
      list(className = 'dt-center', targets = c(0, 2, 3, 4, 5, 6))
    )
  ),
  rownames = FALSE,
  class = 'cell-border stripe hover',
  caption = htmltools::tags$caption(
    style = 'caption-side: top; text-align: left; color: #d97706; font-weight: bold; font-family: Georgia; font-size: 14px;',
    'State-Year Level Alzheimer\'s Cognitive Prevalence Data (2015-2022)'
  )
) |>
  formatStyle(
    'Prevalence (%)',
    background = styleColorBar(range(display_data$`Prevalence (%)`, na.rm = TRUE), '#fcd34d'),
    backgroundSize = '98% 80%',
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'center'
  ) |>
  formatStyle(
    columns = 1:7,
    fontFamily = "Georgia"
  )
```

</div>

<div class="content-section">

# Shiny Dashboard

For a more comprehensive and fully interactive exploration experience, we've created a **Shiny Dashboard** that allows you to:

- **Select specific states** for detailed comparison
- **Choose different time periods** to analyze trends
- **Compare multiple socioeconomic factors** simultaneously
- **Download customized visualizations** and data
- **Explore correlations** between factors interactively

<div class="viz-container">

### Dashboard Features

| Feature | Description |
|---------|-------------|
| **State Selector** | Choose one or multiple states to focus your analysis |
| **Year Range Slider** | Filter data to specific time periods |
| **Factor Comparison** | Toggle between income, education, and air quality views |
| **Dynamic Maps** | Choropleth maps that update based on your selections |
| **Trend Analysis** | Animated time series showing changes over time |
| **Data Export** | Download filtered data in CSV format |

</div>

### Launch the Dashboard Online

Click the button below to access the interactive Shiny dashboard:

<center>
<a href="https://wenye404.shinyapps.io/shiny_app/" target="_blank" 
   style="display: inline-block; padding: 18px 40px; background: linear-gradient(135deg, #d97706 0%, #ea580c 100%); 
          color: white; text-decoration: none; border-radius: 10px; font-weight: bold; font-family: Georgia;
          font-size: 1.1em; box-shadow: 0 4px 15px rgba(217, 119, 6, 0.3); transition: transform 0.2s, box-shadow 0.2s;
          margin: 20px 0;"
   onmouseover="this.style.transform='translateY(-3px)'; this.style.boxShadow='0 8px 25px rgba(217, 119, 6, 0.4)';"
   onmouseout="this.style.transform='translateY(0)'; this.style.boxShadow='0 4px 15px rgba(217, 119, 6, 0.3)';">
  🔍 Launch Interactive Dashboard
</a>
</center>

<div class="viz-container">

### Run Locally (Alternative)

You can also run the dashboard locally on your machine:

**Step 1:** Install required packages:
```r
install.packages(c("shiny", "shinydashboard", "shinyWidgets", 
                   "tidyverse", "plotly", "DT", "viridis", "janitor"))
```

**Step 2:** Launch from R/RStudio:
```r
shiny::runApp("shiny_app")
```

</div>

**Note:** The online dashboard may take a few seconds to load initially.

</div>

<div class="content-section">

# Summary of Findings

Based on our interactive exploration:

1. **Geographic Patterns**: Southern and Southeastern states consistently show higher Alzheimer's-related cognitive prevalence rates compared to Western and Northeastern states.

2. **Temporal Stability**: National average prevalence has remained relatively stable between 25-27% among adults aged 65+ from 2015-2022, with considerable between-state variation.

3. **Demographic Differences**: 
   - Women report higher cognitive burden than men (~31% vs ~27%)
   - Adults aged 50-64 show higher prevalence than those 65+ (~32% vs ~26%)

4. **Socioeconomic Associations**:
   - Higher state-level income is weakly associated with lower prevalence
   - Education shows a modest negative correlation with cognitive burden
   - Air quality (PM2.5) shows minimal association at the ecological level

5. **State Variability**: Individual states can vary by 10+ percentage points in prevalence, suggesting important local factors beyond what aggregate socioeconomic indicators capture.

</div>